---
title: "ST117 E2"
author: "Homework Lab Group 016 Pod C"
output:
  pdf_document: default
  html_document: default
---

-   Please read the document "Guidance on Exercise Sets" on Moodle.
-   Please replace xyz and X in the header by the correct codes for your Homework Pod.
-   Please list all Homework Pod members who contributed to these solutions and which questions/parts they are responsible for (final write up).
-   We put together some generic examples for TEX commands, inclusion of photos or images, and for data visualisation. See at the end of the document. Please delete them before submission.
-   See there also for guidance about R code chunks vs rest of document.
-   Please remember to submit both this Rmd file AND a the pdf output after knitting.

# This submission was created by:

1.  Name and WARWICK ID: Liuyinuo Du, 5694901: A 1
2.  Name and WARWICK ID: Chigozie Iheanyichukwu, 2259977: A 2,3
3.  Name and WARWICK ID: Lichangxuan Qb, 5695200: B 1,2
4.  Name and WARWICK ID: Lopich Jan, 5678772: B 3

```{r setup, include=FALSE}
library(knitr)
```

# Question A: Covariance and correlation

## 1. Theory

### 1.(a)

```{r}
set.seed(123)  

X <- rnorm(10000)   
Y <- rnorm(10000)

a <- 5             
b <- 3              

const_a <- rep(a, length(X))
cov_X_a <- cov(X, const_a)
cov_X_a   
left_side <- cov(a*X + b, Y)      
right_side <- a * cov(X, Y)       
```

### 1.(b)

```{r}
set.seed(123)

X <- rnorm(10000)
Y <- rnorm(10000)
U <- rnorm(10000)

a <- 2
b <- -4

left_side <- cov(a*X + b*Y, U)
right_side <- a*cov(X,U) + b*cov(Y,U)
```

### 1.(c)

```{r}
set.seed(123)

X <- rnorm(10000)
U <- rnorm(10000)
V <- rnorm(10000)

a <- 3
b <- 7

left_side <- cov(X, a*U + b*V)
right_side <- a*cov(X,U) + b*cov(X,V)
```

### 1.(d)

```{r}
set.seed(123)

X <- rnorm(10000)
Y <- rnorm(10000)
cov(X,Y)   

set.seed(123)

X <- rnorm(10000)
Y <- rnorm(10000)

cov(X,Y)   
```

### 1.(e)

```{r}
set.seed(123)

X1 <- rnorm(10000)
X2 <- rnorm(10000)
X3 <- rnorm(10000)

S <- X1 + X2 + X3
left_side <- var(S)
right_side <- var(X1) + var(X2) + var(X3) +
  2*cov(X1,X2) +
  2*cov(X1,X3) +
  2*cov(X2,X3)
```

### 1.(f)

```{r}
set.seed(123)

X <- rnorm(10000)
Y <- 0.8*X + rnorm(10000, sd=0.2)  

rho <- cor(X,Y)
Y_neg <- -X + rnorm(10000, sd=0.1)

cor(X, Y_neg)
```

## 2. Examples
### 2.(a)

Cov(X,Y)= E[XY]-E[X]E[Y]
        =  0n+(1/36)n(n-1)=(n(n-1))/36 - (n^2)/36
        =  -n/36     <- A)
        
Var(X), Var(Y) = 5n/36

(-n/36)/sqrt((5n/36)^2) = -1/5   <-  B)

### 2.(b)

Cov(Z,W) = E[ZW] - E[Z]E[W]  
          = 1/4 - (2/3)(1/3) = 1/4 - 2/9 = 1/36    <- A)

Var(Z) = 1/2 - (2/3)^2 = 1/18, and by symmetry Var(W) = 1/18

(1/36) / sqrt((1/18)^2) = (1/36)/(1/18) = 1/2    <- B)

## 3. Simulations
### 3.(a)
```{r}
set.seed(42)
n <- 200

get_data <- function(corr, n) {
  x <- rnorm(n)
  y <- corr * x + sqrt(1 - corr^2) * rnorm(n)
  list(x=x, y=y)
}


cors_ <- c(-1, 0, 0.5, 0.9)
for (rho in cors_) {
  d <- get_data(rho, n)
  actual <- round(cor(d$x, d$y), 3)
  plot(d$x, d$y,
       main = paste0("p = ", rho),
       xlab = "X", ylab = "Y")
}
```

### 3.(b)

***Why r is ~0 but the variables are dependent on each other?

r measured in this way (Pearson's correlation), only measures linear association. The curve Y=X^2 is obviously not linear, and positive deviations on the left side of the curve will be cancelled out by negative deviations on the right side of the curve. Making the numerator close to 0, and r very small.
```{r}

x = seq(-3, 3, length.out = 300)
y = x^2 + rnorm(300, sd = 0.3)

actual <- round(cor(x, y), 3)
plot(x, y,
     main = paste0("Y = X^2  | r = ", actual),
     xlab = "X", ylab = "Y",
     pch =19, cex= 0.6)
```

# Question B: Sampling

## 1. Theory

The population is $\{1, 2, 2, 4, 8\}$ with $N = 5$.

**Population mean:**

$$\mu = \frac{1+2+2+4+8}{5} = \frac{17}{5} = 3.4$$

**Population variance:**

$$
\sigma^2 = \frac{(1-3.4)^2+(2-3.4)^2+(2-3.4)^2+(4-3.4)^2+(8-3.4)^2}{5}
= \frac{5.76+1.96+1.96+0.36+21.16}{5} = \frac{31.2}{5} = 6.24
$$

### 1.(a)

Drawing samples of size $n = 2$ from $N = 5$ elements yields $\binom{5}{2} = 10$ equally likely samples, each with probability $\frac{1}{10}$.

To distinguish the two 2's, label them $2_a$ and $2_b$.

All possible samples and their means:

|     Sample      | Sample Mean $\bar{x}$ |
|:---------------:|:---------------------:|
|  $\{1,\ 2_a\}$  |          1.5          |
|  $\{1,\ 2_b\}$  |          1.5          |
|   $\{1,\ 4\}$   |          2.5          |
|   $\{1,\ 8\}$   |          4.5          |
| $\{2_a,\ 2_b\}$ |          2.0          |
|  $\{2_a,\ 4\}$  |          3.0          |
|  $\{2_a,\ 8\}$  |          5.0          |
|  $\{2_b,\ 4\}$  |          3.0          |
|  $\{2_b,\ 8\}$  |          5.0          |
|   $\{4,\ 8\}$   |          6.0          |

Sampling distribution of $\bar{X}$:

| $\bar{x}$ | 1.5 | 2.0 | 2.5 | 3.0 | 4.5 | 5.0 | 6.0 |
|:-------:|:-------:|:-------:|:-------:|:-------:|:-------:|:-------:|:-------:|
| $P(\bar{X}=\bar{x})$ | $\frac{2}{10}$ | $\frac{1}{10}$ | $\frac{1}{10}$ | $\frac{2}{10}$ | $\frac{1}{10}$ | $\frac{2}{10}$ | $\frac{1}{10}$ |

Mean of $\bar{X}$:

$$E[\bar{X}] = 1.5\cdot\frac{2}{10}+2\cdot\frac{1}{10}+2.5\cdot\frac{1}{10}+3\cdot\frac{2}{10}
+4.5\cdot\frac{1}{10}+5\cdot\frac{2}{10}+6\cdot\frac{1}{10}
= \frac{34}{10} = 3.4$$

Variance of $\bar{X}$:

$$E[\bar{X}^2] = \frac{1.5^2\cdot2 + 2^2\cdot1 + 2.5^2\cdot1 + 3^2\cdot2 + 4.5^2\cdot1
+ 5^2\cdot2 + 6^2\cdot1}{10} = \frac{139}{10} = 13.9$$

$$\text{Var}(\bar{X}) = E[\bar{X}^2] - \left(E[\bar{X}]\right)^2 = 13.9 - 3.4^2 = 13.9 - 11.56
= 2.34$$

Theoretical check using the Finite Population Correction (FPC):

$$\text{Var}(\bar{X}) = \frac{\sigma^2}{n}\cdot\frac{N-n}{N-1}
= \frac{6.24}{2}\cdot\frac{5-2}{5-1} = 3.12 \times 0.75 = 2.34 \quad$$

### 1.(b)

With replacement, there are $5^2 = 25$ ordered pairs, each with probability $\frac{1}{25}$.

Sampling distribution of $\bar{X}$:

| $\bar{x}$ | 1.0 | 1.5 | 2.0 | 2.5 | 3.0 | 4.0 | 4.5 | 5.0 | 6.0 | 8.0 |
|:-------:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|
| Count | 1 | 4 | 4 | 2 | 4 | 1 | 2 | 4 | 2 | 1 |
| $P(\bar{X}=\bar{x})$ | $\frac{1}{25}$ | $\frac{4}{25}$ | $\frac{4}{25}$ | $\frac{2}{25}$ | $\frac{4}{25}$ | $\frac{1}{25}$ | $\frac{2}{25}$ | $\frac{4}{25}$ | $\frac{2}{25}$ | $\frac{1}{25}$ |

Mean of $\bar{X}$:

$$E[\bar{X}] = \frac{1(1)+1.5(4)+2(4)+2.5(2)+3(4)+4(1)+4.5(2)+5(4)+6(2)+8(1)}{25}
= \frac{85}{25} = 3.4$$

Variance of $\bar{X}$:

$$E[\bar{X}^2] = \frac{1(1)+2.25(4)+4(4)+6.25(2)+9(4)+16(1)+20.25(2)+25(4)+36(2)+64(1)}{25}
= \frac{367}{25} = 14.68$$

$$\text{Var}(\bar{X}) = 14.68 - 3.4^2 = 14.68 - 11.56 = 3.12$$

Theoretical check:

$$\text{Var}(\bar{X}) = \frac{\sigma^2}{n} = \frac{6.24}{2} = 3.12 \quad$$

**Comparison and Discussion**

|                       | Without Replacement | With Replacement |
|:----------------------|:-------------------:|:----------------:|
| $E[\bar{X}]$          |         3.4         |       3.4        |
| $\text{Var}(\bar{X})$ |        2.34         |       3.12       |

**Mean:** Both methods yield $E[\bar{X}] = \mu = 3.4$, confirming that the sample mean is an **unbiased estimator** of the population mean regardless of the sampling scheme.

**Variance:** The two methods differ because sampling without replacement benefits from the **Finite Population Correction (FPC)**:

$$\text{Var}_{\text{without}}(\bar{X}) = \frac{\sigma^2}{n}\cdot\underbrace{\frac{N-n}{N-1}}_{\text{FPC}=0.75}
< \frac{\sigma^2}{n} = \text{Var}_{\text{with}}(\bar{X})$$

When sampling without replacement, each draw is guaranteed to provide a distinct observation, eliminating the possibility of selecting the same unit twice. This results in less variability in the sample mean, making it a more precise estimator. As $N \to \infty$, the FPC approaches 1 and the two variances converge.

## 2. Simulations

```{r}
set.seed(42)

# 1. Construct the finite population
# Value k appears k times, so pop = {1,1, 2,2,2, 3,3,3,3, ..., 20x20}
pop      <- rep(1:20, 1:20)
N        <- length(pop)                    # total population size = 210
pop_mean <- mean(pop)                      # population mean
pop_var  <- mean((pop - pop_mean)^2)       # population variance (divided by N)

cat("Population size  N  =", N, "\n")
cat("Population mean  mu  =", round(pop_mean, 4), "\n")
cat("Population var   sigma^2 =", round(pop_var,  4), "\n\n")


# 2. Simulation function
simulate_sampling <- function(n, replace, B = 1000) {

  # Draw B samples of size n and record each sample mean
  sample_means <- replicate(B, mean(sample(pop, n, replace = replace)))

  # Empirical (simulated) statistics
  emp_mean <- mean(sample_means)
  emp_var  <- var(sample_means)

  # Theoretical statistics
  th_mean <- pop_mean
  if (replace) {
    th_var <- pop_var / n                              # standard result
  } else {
    th_var <- (pop_var / n) * (N - n) / (N - 1)       # finite population correction (FPC)
  }

  # Print results
  method_label <- if (replace) "With replacement (b)" else "Without replacement (a)"
  cat(sprintf("  [%s]  n = %d\n", method_label, n))
  cat(sprintf("    Empirical  mean = %.4f  |  var = %.6f\n", emp_mean, emp_var))
  cat(sprintf("    Theoretical mean = %.4f  |  var = %.6f\n", th_mean, th_var))
  if (!replace) {
    cat(sprintf("    FPC = (N-n)/(N-1) = (%d-%d)/(%d-1) = %.4f\n",
                N, n, N, (N - n) / (N - 1)))
  }
  cat("\n")

  invisible(sample_means)   # return means silently for optional further use
}
```

### 2.(a)

```{r}
# Part (a): Sampling WITHOUT replacement
means_a_10  <- simulate_sampling(n = 10,  replace = FALSE)
means_a_200 <- simulate_sampling(n = 200, replace = FALSE)
```

### 2.(b)

```{r}
# Part (b): Sampling WITH replacement
means_b_10  <- simulate_sampling(n = 10,  replace = TRUE)
means_b_200 <- simulate_sampling(n = 200, replace = TRUE)
```

For **n = 10**, (a) and (b) give nearly identical variances because the FPC $\approx$ 0.957 is close to 1. removing 10 items from a population of 210 barely changes the probabilities.

For **n = 200**, the difference is dramatic. Without replacement the variance collapses (FPC $\approx$ 0.048) because 200 out of 210 elements are observed, leaving almost no uncertainty. With replacement the variance stays large since repeated draws are allowed and the same information gain is not guaranteed.

Both methods yield empirical means very close to the true mu in all four cases, confirming that the sample mean is an **unbiased estimator** of the population mean regardless of sampling scheme or sample size.

## 3. Sampling fish
### 3.(a) 

$$
\begin{aligned}
P(S<10) &= P(S \leq 9) = P(S=0) + P(S=1) + \ldots + P(S=9) \\
\\
P(S=0) &= (4/20)^{15} \\
P(S=1) &= \frac{16}{20} \times (4/19)^{14} + \frac{4}{20} \times \frac{16}{20} \times (4/19)^{13} + \text{... all other combinations} \\
P(S=2) &= \text{all combinations of 13 small \& 2 big fish} \\
&\vdots \\
P(S=9) &= \text{all combinations of 9 big \& 6 small fish} \\
\\
&\text{Sum the probabilities together to get } P(S<10)
\end{aligned}
$$

```{r}

# Function to compute P(S = s) for a given s
prob_S_equals_s <- function(s) {
  if (s > 15 || s < 0) return(0)
  
  positions <- combn(15, s, simplify = FALSE)
  
  total_prob <- 0
  
  for (pos in positions) {
    
    p <- 1
    bigFishCaught <- 0  
    
    for (attempt in 1:15) {
      bigInPond <- 16 - bigFishCaught
      smallInPond <- 4
      total <- bigInPond + smallInPond
      
      if (attempt %in% pos) {
        p <- p * (bigInPond / total)
        bigFishCaught <- bigFishCaught + 1
      }
      
      else {
        p <- p * (smallInPond / total)
      }
    }
    
    total_prob <- total_prob + p
  }
  
  return(total_prob)
}

# Function to compute P(S <= 9)
probS <- function() {
  total <- 0
  for (s in 0:9) {
    p_s <- prob_S_equals_s(s)
    cat(sprintf("P(S = %d) = %.8f\n", s, p_s))
    total <- total + p_s
  }
  return(total)
}

# Results
result <- probS()
cat(sprintf("\nP(S < 10) = %.8f\n", result))

```


Initially, p = 0.8 = 16/20 
Since p gets smaller as big fish get caught, an upper bound would be to model the fish as X~Bin(15, 0.8) 

Therefore, an upper bound would be:
```{r}

u_bound <- pbinom(9, 15, 0.8)
cat(sprintf("Upper bound for P(S < 10) <= %.8f", u_bound))

```

We know that the expected mean for sampling without replacement is simply the population mean = 625g 

We also need to take into account that S=0 is a possibility. However, its value is 0.2^15 = 3.0517578e-11 

So the overall expected mean would be 625 * (1 - 3.0517578e-11) ~ 625

### 3.(b)

```{r} 

S_values = c()
means = c()

for (s in 1:1000) {
  fishpond <- seq(50, 1000, 50)
  sampledFish <- c()
  iterations <- 0
  bigFish <- 0 
  
  while (bigFish < 10 && iterations < 15) { 
    iterations <- iterations + 1 
    
    l = length(fishpond) 
    ind = sample(1:l, 1) 
    
    fishSize <- fishpond[ind] 
    if (fishSize > 200) {
      fishpond <- fishpond[fishpond != fishSize] 
      sampledFish = append(sampledFish, fishSize)
      bigFish <- bigFish + 1
    }

  } 
  
  S_values <- append(S_values, length(sampledFish))
  means <- append(means, mean(sampledFish))
  
} 

# Results
cat("P(S < 10):", sum(S_values < 10) / 1000, "\n")
cat("E[Xbar]:", mean(means), "\n")

```  
