---
title: "ST117 E3"
author: "Homework Lab Group xyz Pod X"
output:
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

-   Please read the document "Guidance on Exercise Sets" on Moodle.
-   Please replace xyz and X in the header by the correct codes for your
    Homework Pod.
-   Please list all Homework Pod members who contributed to these
    solutions and which questions/parts they are responsible for (final
    write up).
-   We put together some generic examples for TEX commands, inclusion of
    photos or images, and for data visualisation. See at the end of the
    document. Please delete them before submission.
-   Please use R code chunks only for code and very brief technical
    comments tied directly to specific lines. All other text (e.g.
    explanations of coding, statistical reasoning, analytical steps, and
    conclusions) should be written outside the chunks.
-   Please remember to submit both this Rmd file AND a the pdf output
    after knitting.

# This submission was created by:

1.  Name and WARWICK ID: Chigozie Iheanyichukwu 2259977: A 1,2,3
2.  Name and WARWICK ID: Liuyinuo Du 5694901: B 1,2 
3.  Name and WARWICK ID: Changxuan Li 5695200: B 3,4
4.  Name and WARWICK ID: Lopich Jan 5678772: B 5,6

```{r setup, include=FALSE}
#Code to setup packages should go here, for example:
library(knitr)
```

# Question A: Thief in the kitchen

## 1. Derive MoM and MLE estimators

$X_1, ..., X_n$ represent i.i.d random variables representing waiting times in days between thefts. The waiting time includes the day the theft happens so $X_i \geq 1$. We model them as geometric as per the question:

$$P(X = k) = (1-p)^{k-1} p, \quad where \  k = 1, 2, 3, ...$$

The mean of the geometric distribution is $E[X] = 1/p$.

We set the first population moment equal to the first sample moment i.e $\frac{1}{p} = \bar{X}$$

so **for MoM** solving for p gives us:

$$\hat{p} = \frac{1}{\bar{X}} = \frac{n}{\sum_{i=1}^{n} X_i}$$

For MLE, The use the log-likelihood for n observations.

$$l(p) = n \log(p) + \left(\sum_{i=1}^{n} x_i - n \right) \log(1-p)$$

Then we differentiate and equal to 0:

$$\frac{dl}{dp} = \frac{n}{p} - \frac{\sum x_i - n}{1-p} = 0$$

So $n(1-p) = p(\sum x_i - n)$, and $n = p \sum x_i$ Hence **for MLE**, p is:

$$\hat{p}_{mle} = \frac{n}{\sum_{i=1}^{n} X_i} = \frac{1}{\bar{X}}$$

The second derivative is $-n/p^2 - (\sum x_i - n)/(1-p)^2 < 0$ so this is a maximum. Interestingly, MoM and MLE give the same estimator for the geometric distribution.

## 2. Bias?

We want to check if $E[\hat{p}] = p$. Our estimator is $\hat{p} = 1/\bar{X}$.

The function $h(x) = 1/x$ has second derivative $h''(x) = 2/x^3 > 0$ for $x > 0$, so it is convex. as per Jensens inequality we get:

$$E[\hat{p}] = E\left[\frac{1}{\bar{X}}\right] > \frac{1}{1/p} = p$$

So $E[\hat{p}] > p$ which means the estimator is biased upwards and tends to overestimate p. (MoM and MLE estimator are biased since they are the same thing)

## 3. Fitting using dataset

```{r}
# read in the data
thefts <- scan("foodTheftsE3.csv")
cat("Number of days:", length(thefts), "\n")
cat("Days with theft:", sum(thefts), "\n")
```

```{r}
# find which days had thefts
theft_days <- which(thefts == 1)
cat("Theft occurred on days:", theft_days, "\n")
```

```{r}
# compute waiting times between consecutive thefts
waiting_times <- diff(theft_days)
cat("Waiting times:", waiting_times, "\n")
n <- length(waiting_times)
cat("Number of waiting times (n):", n, "\n")
```

```{r}
# compute the estimators
xbar <- mean(waiting_times)
cat("Sample mean:", xbar, "\n")

p_hat <- 1 / xbar
cat("MoM estimate of p:", p_hat, "\n")
cat("MLE estimate of p:", p_hat, "\n")
```

We have $n = 17$ waiting times between the 18 thefts. The sample mean is $\bar{x} \approx 2.41$ days, so our estimate is

$$\hat{p}_{MoM} = \hat{p}_{MLE} = \frac{1}{\bar{x}} = \frac{1}{2.41} \approx 0.415$$

So we estimate roughly a 41.5% chance of a theft on any given day. Though, this is a s

# Question B: Analysis of exam score distributions

## 1. Finding NAs

```{r} 
Exams <- read.csv("syntheticExamScoresE3.csv")

Ex118_NA <- is.na(Exams$Ex118)
ExMA_NA  <- is.na(Exams$ExMA)
NA_table <- table(ST118_NA = Ex118_NA,
                  Maths_NA = ExMA_NA)

NA_table
addmargins(NA_table)

both_taken <- sum(!Ex118_NA & !ExMA_NA)

only_118 <- sum(!Ex118_NA & ExMA_NA)

only_MA <- sum(Ex118_NA & !ExMA_NA)

neither <- sum(Ex118_NA & ExMA_NA)

both_taken
only_118
only_MA
neither
```

## 2. Descriptive analysis

```{r}
Ex118_clean <- Exams$Ex118[!is.na(Exams$Ex118)]

mean(Ex118_clean)

sd(Ex118_clean)

median(Ex118_clean)

min(Ex118_clean)

max(Ex118_clean)

quantile(Ex118_clean)

ExMA_clean <- Exams$ExMA[!is.na(Exams$ExMA)]

mean(ExMA_clean)

sd(ExMA_clean)

median(ExMA_clean)

min(ExMA_clean)

max(ExMA_clean)

quantile(ExMA_clean)

hist(Ex118_clean,
     breaks = 20,
     probability = TRUE,
     main = "Histogram of ST118 Scores",
     xlab = "ST118 Score")

hist(ExMA_clean,
     breaks = 20,
     probability = TRUE,
     main = "Histogram of Maths Scores",
     xlab = "Maths Score")

complete_data <- Exams[complete.cases(Exams$Ex118, Exams$ExMA), ]

head(complete_data)

plot(complete_data$Ex118,
     complete_data$ExMA,
     main = "Scatterplot of ST118 vs Maths",
     xlab = "ST118 Score",
     ylab = "Maths Score")

cor(complete_data$Ex118,
    complete_data$ExMA)

plot(jitter(complete_data$Ex118),
     jitter(complete_data$ExMA),
     main = "Scatterplot with Jitter",
     xlab = "ST118 Score",
     ylab = "Maths Score")

plot(complete_data$Ex118,
     complete_data$ExMA,
     col = rgb(0,0,0,0.4),
     pch = 16,
     main = "Scatterplot with Transparency",
     xlab = "ST118 Score",
     ylab = "Maths Score")

pairs <- paste(complete_data$Ex118,
               complete_data$ExMA)

pair_counts <- table(pairs)

duplicated_pairs <- pair_counts[pair_counts > 1]

students_in_duplicates <- sum(duplicated_pairs)

students_in_duplicates
```

## 3. MoM for Normal

```{r}
Exams <- read.csv("syntheticExamScoresE3.csv")

# Extract data by group
x118_all <- na.omit(Exams$Ex118)          # including zeros
xMA_all  <- na.omit(Exams$ExMA)
x118_pos <- x118_all[x118_all > 0]       # excluding zeros
xMA_pos  <- xMA_all[xMA_all > 0]
```

```{r}
# MOM estimators = sample mean + sample standard deviation
mu118  <- mean(x118_all);  sig118  <- sd(x118_all)
muMA   <- mean(xMA_all);   sigMA   <- sd(xMA_all)
mu118p <- mean(x118_pos);  sig118p <- sd(x118_pos)
muMAp  <- mean(xMA_pos);   sigMAp  <- sd(xMA_pos)

cat("Normal Distribution Parameters (including zeros) \n")
cat("Ex118: mu =", round(mu118,4),  "sigma =", round(sig118,4),  "\n")
cat("ExMA:  mu =", round(muMA,4),   "sigma =", round(sigMA,4),   "\n")
cat("Normal Distribution Parameters (excluding zeros) \n")
cat("Ex118: mu =", round(mu118p,4), "sigma =", round(sig118p,4), "\n")
cat("ExMA:  mu =", round(muMAp,4),  "sigma =", round(sigMAp,4),  "\n")
```

```{r}
# Histogram + fitted normal curve
par(mfrow=c(2,2))
hist(x118_all, probability=TRUE, breaks=20, col="lightblue",
     main="Ex118 (incl. zeros) — Normal Fit", xlab="Score")
curve(dnorm(x, mu118, sig118), add=TRUE, col="red", lwd=2)

hist(xMA_all, probability=TRUE, breaks=20, col="lightgreen",
     main="ExMA (incl. zeros) — Normal Fit", xlab="Score")
curve(dnorm(x, muMA, sigMA), add=TRUE, col="red", lwd=2)

hist(x118_pos, probability=TRUE, breaks=20, col="lightblue",
     main="Ex118 (excl. zeros) — Normal Fit", xlab="Score")
curve(dnorm(x, mu118p, sig118p), add=TRUE, col="red", lwd=2)

hist(xMA_pos, probability=TRUE, breaks=20, col="lightgreen",
     main="ExMA (excl. zeros) — Normal Fit", xlab="Score")
curve(dnorm(x, muMAp, sigMAp), add=TRUE, col="red", lwd=2)
```

```{r}
# Q-Q plots
par(mfrow=c(2,2))
qqnorm(x118_all, main="Q-Q Plot: Ex118 (incl. zeros)");  qqline(x118_all, col="red", lwd=2)
qqnorm(xMA_all,  main="Q-Q Plot: ExMA (incl. zeros)");   qqline(xMA_all,  col="red", lwd=2)
qqnorm(x118_pos, main="Q-Q Plot: Ex118 (excl. zeros)");  qqline(x118_pos, col="red", lwd=2)
qqnorm(xMA_pos,  main="Q-Q Plot: ExMA (excl. zeros)");   qqline(xMA_pos,  col="red", lwd=2)
```

## 4. Parameter calculations and MoM for Gamma

#### 1. The Gamma Distribution

A random variable $X \sim \text{Gamma}(\alpha, \beta)$ with shape
$\alpha > 0$ and rate $\beta > 0$ has p.d.f.:

$$
f(x;\,\alpha,\beta)
  = \frac{\beta^{\alpha}}{\Gamma(\alpha)}\, x^{\alpha - 1}\, e^{-\beta x},
  \qquad x > 0,
$$

where $\Gamma(\alpha) = \int_0^{\infty} t^{\alpha-1} e^{-t}\, dt$
satisfies the recurrence $\Gamma(\alpha + 1) = \alpha\,\Gamma(\alpha)$.

#### 2. Mean

$$
\mathbb{E}[X]
  = \int_0^{\infty} x \cdot \frac{\beta^{\alpha}}{\Gamma(\alpha)}\,
    x^{\alpha-1}\, e^{-\beta x}\, dx
  = \frac{\beta^{\alpha}}{\Gamma(\alpha)}
    \int_0^{\infty} x^{\alpha}\, e^{-\beta x}\, dx.
$$

**Substitution** $u = \beta x$ $\Rightarrow$ $x = u/\beta$,
$dx = du/\beta$:

$$
\int_0^{\infty} x^{\alpha}\, e^{-\beta x}\, dx
  = \frac{1}{\beta^{\alpha+1}} \int_0^{\infty} u^{\alpha}\, e^{-u}\, du
  = \frac{\Gamma(\alpha + 1)}{\beta^{\alpha+1}}.
$$

Substituting back and using $\Gamma(\alpha+1) = \alpha\,\Gamma(\alpha)$:

$$
\begin{aligned}
\mathbb{E}[X]
&= \frac{\beta^{\alpha}}{\Gamma(\alpha)} \cdot
   \frac{\Gamma(\alpha+1)}{\beta^{\alpha+1}} \\
&= \frac{\alpha\,\Gamma(\alpha)}{\beta\,\Gamma(\alpha)} \\
&= \frac{\alpha}{\beta}.
\end{aligned}
$$

#### 3. Variance

Given $n$ i.i.d. observations $x_1,\ldots,x_n > 0$, define:

$$
\bar{x} = \frac{1}{n}\sum_{i=1}^n x_i,
\qquad
m_2 = \frac{1}{n}\sum_{i=1}^n (x_i - \bar{x})^2.
$$

> $m_2$ divides by $n$ (not $n-1$) because MoM equates population
> moments directly to sample moments.

**System of equations** (population = sample):

$$
\begin{cases}
\dfrac{\alpha}{\beta} = \bar{x} \\[8pt]
\dfrac{\alpha}{\beta^2} = m_2
\end{cases}
$$

**Solving:** divide the first equation by the second:

$$
\frac{\alpha/\beta}{\alpha/\beta^2} = \beta = \frac{\bar{x}}{m_2}
\quad\Longrightarrow\quad
\hat{\beta} = \frac{\bar{x}}{m_2}.
$$

Substitute back into the first equation:

$$
\hat{\alpha} = \hat{\beta}\,\bar{x} = \frac{\bar{x}^2}{m_2}.
$$

The **MoM estimators** for the Gamma distribution are:

$$
  \hat{\alpha}_{\text{MoM}} = \frac{\bar{x}^2}{m_2},
  \qquad
  \hat{\beta}_{\text{MoM}}  = \frac{\bar{x}}{m_2}.
$$

##### 3.1 Second Moment

$$
\mathbb{E}[X^2]
  = \frac{\beta^{\alpha}}{\Gamma(\alpha)}
    \int_0^{\infty} x^{\alpha+1}\, e^{-\beta x}\, dx
  = \frac{\beta^{\alpha}}{\Gamma(\alpha)} \cdot
    \frac{\Gamma(\alpha+2)}{\beta^{\alpha+2}}.
$$

Applying the recurrence twice,
$\Gamma(\alpha+2) = (\alpha+1)\alpha\,\Gamma(\alpha)$:

$$
\mathbb{E}[X^2] = \frac{\alpha(\alpha+1)}{\beta^2}.
$$

##### 3.2 Variance

$$
\begin{aligned}
\operatorname{Var}(X)
&= \mathbb{E}[X^2] - (\mathbb{E}[X])^2 \\
&= \frac{\alpha(\alpha+1)}{\beta^2} - \frac{\alpha^2}{\beta^2} \\
&= \frac{\alpha^2 + \alpha - \alpha^2}{\beta^2} \\
&= \frac{\alpha}{\beta^2}.
\end{aligned}
$$

#### 4. Method of Moments Estimators

Given $n$ i.i.d. observations $x_1,\ldots,x_n > 0$, define:

$$
\bar{x} = \frac{1}{n}\sum_{i=1}^n x_i,
\qquad
m_2 = \frac{1}{n}\sum_{i=1}^n (x_i - \bar{x})^2.
$$

$m_2$ divides by $n$ (not $n-1$) because MoM equates population \>
moments directly to sample moments.

**System of equations** (population = sample):

$$
\begin{cases}
\dfrac{\alpha}{\beta} = \bar{x} \\[8pt]
\dfrac{\alpha}{\beta^2} = m_2
\end{cases}
$$

**Solving:** divide the first equation by the second:

$$
\frac{\alpha/\beta}{\alpha/\beta^2} = \beta = \frac{\bar{x}}{m_2}
\quad\Longrightarrow\quad
\hat{\beta} = \frac{\bar{x}}{m_2}.
$$

Substitute back into the first equation:

$$
\hat{\alpha} = \hat{\beta}\,\bar{x} = \frac{\bar{x}^2}{m_2}.
$$

The **MoM estimators** for the Gamma distribution are:

$$
  \hat{\alpha}_{\text{MoM}} = \frac{\bar{x}^2}{m_2},
  \qquad
  \hat{\beta}_{\text{MoM}}  = \frac{\bar{x}}{m_2}.
$$

```{r}
mom_gamma <- function(x) {
  xbar      <- mean(x)
  s2        <- var(x)
  alpha_hat <- xbar^2 / s2
  beta_hat  <- xbar  / s2
  return(c(alpha = alpha_hat, beta = beta_hat))
}

cat("Gamma MoM Parameters (excluding zeros)\n")
cat("Ex118: "); g118 <- mom_gamma(x118_pos)
cat("ExMA:  "); gMA  <- mom_gamma(xMA_pos)

# Histograms + Fitted Gamma Density

par(mfrow = c(1, 2))

hist(x118_pos, probability = TRUE, breaks = 20, col = "lightblue",
     main = "Ex118 (excl. zeros) — Gamma Fit", xlab = "Score")
curve(dgamma(x, shape = g118["alpha"], rate = g118["beta"]),
      add = TRUE, col = "darkblue", lwd = 2)
legend("topleft",
       legend = paste0("alpha=", round(g118["alpha"],3),
                       "  beta=", round(g118["beta"],3)),
       col = "darkblue", lwd = 2, bty = "n", cex = 0.85)

hist(xMA_pos, probability = TRUE, breaks = 20, col = "lightgreen",
     main = "ExMA (excl. zeros) — Gamma Fit", xlab = "Score")
curve(dgamma(x, shape = gMA["alpha"], rate = gMA["beta"]),
      add = TRUE, col = "darkgreen", lwd = 2)
legend("topleft",
       legend = paste0("alpha=", round(gMA["alpha"],3),
                       "  beta=", round(gMA["beta"],3)),
       col = "darkgreen", lwd = 2, bty = "n", cex = 0.85)


# Gamma Q-Q plots (theoretical vs empirical quantiles)
# Points on the y=x line => good fit

par(mfrow = c(1, 2))

# Ex118
p118 <- ppoints(length(x118_pos))
q_th_118 <- qgamma(p118, shape = g118["alpha"], rate = g118["beta"])
q_em_118 <- sort(x118_pos)
plot(q_th_118, q_em_118,
     main = "Gamma Q-Q: Ex118 (excl. zeros)",
     xlab = "Theoretical Quantiles (Gamma)",
     ylab = "Sample Quantiles",
     pch = 16, cex = 0.6, col = "darkblue")
abline(0, 1, col = "red", lwd = 2)

# ExMA
pMA  <- ppoints(length(xMA_pos))
q_th_MA <- qgamma(pMA, shape = gMA["alpha"], rate = gMA["beta"])
q_em_MA <- sort(xMA_pos)
plot(q_th_MA, q_em_MA,
     main = "Gamma Q-Q: ExMA (excl. zeros)",
     xlab = "Theoretical Quantiles (Gamma)",
     ylab = "Sample Quantiles",
     pch = 16, cex = 0.6, col = "darkgreen")
abline(0, 1, col = "red", lwd = 2)
```

Based on the output values, the mean and variance are identical to the
sample, indicating that the theory aligns with the data. Compared to the
normal distribution, the Gamma distribution only takes positive values
and can be left-skewed, right-skewed, or even nearly symmetrical. As the
Gamma distribution has no upper bound, it cannot perfectly describe
examination scores constrained within [0, 1], thus potentially
necessitating the use of the Beta distribution.

## 5. Parameter calculations and MoM for Beta

#### Parameter Derivation 

Let $X \sim Beta(\alpha,\beta)$ with pdf

$$
f(x)=\frac{1}{B(\alpha,\beta)} x^{\alpha-1}(1-x)^{\beta-1}, \quad 0<x<1
$$

The first moment is given by:

$$
E[X] = \int_0^1 x f(x) dx
$$

$$
= \frac{1}{B(\alpha,\beta)} \int_0^1 x^{\alpha}(1-x)^{\beta-1} dx
$$

$$
= \frac{B(\alpha+1,\beta)}{B(\alpha,\beta)}
$$

Using $B(a,b)=\frac{\Gamma(a)\Gamma(b)}{\Gamma(a+b)}$, this simplifies to:

$$
E[X] = \frac{\alpha}{\alpha+\beta}
$$

Second moment is given by:

$$
E[X^2] = \int_0^1 x^2 f(x) dx
$$

$$
= \frac{1}{B(\alpha,\beta)} \int_0^1 x^{\alpha+1}(1-x)^{\beta-1} dx
$$

$$
= \frac{B(\alpha+2,\beta)}{B(\alpha,\beta)}
$$
Which through cancellation simplifies to: 

$$
E[X^2] =
\frac{\alpha(\alpha+1)}{(\alpha+\beta)(\alpha+\beta+1)}
$$
Therefore the variance can be calculated through the following: 

$$
\sigma^2 =
E[X^2] - (E[X])^2=
\frac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}
$$

By then dividing the variance by $\bar{x}$ and substituting $\beta = \frac{\alpha(1-\bar{x})}{\alpha}$, you get: 

$$
\hat{\alpha} = 
\left(\frac{(1-\bar{x})\bar{x}}{\sigma^2} - 1\right)\bar{x}
$$
and 

$$
\hat{\beta} = 
\left(\frac{(1-\bar{x})\bar{x}}{\sigma^2} - 1\right)(1-\bar{x})
$$


```{r}
Exams <- read.csv("syntheticExamScoresE3.csv")

# Extract data by group
x118_all <- na.omit(Exams$Ex118)                      # including zeros
xMA_all  <- na.omit(Exams$ExMA)
x118_pos <- x118_all[1 > x118_all & x118_all > 0]     # excluding 0s and 1s
xMA_pos  <- xMA_all[1 > xMA_all & xMA_all > 0]
```


```{r} 

mom_beta <- function(x) {
  xbar      <- mean(x)
  s2        <- var(x)
  alpha_hat <- xbar * ((xbar*(1-xbar)/s2) - 1)
  beta_hat  <- (1-xbar) * ((xbar*(1-xbar)/s2) - 1)
  return(c(alpha = alpha_hat, beta = beta_hat))
}

cat("Gamma MoM Parameters (excluding zeros)\n")
cat("Ex118: "); b_118 <- mom_beta(x118_pos)
cat("ExMA:  "); b_MA  <- mom_beta(xMA_pos) 

# Ex118 Plot
hist(x118_pos, 
     probability = TRUE, 
     breaks = 20, 
     col = "lightblue",
     main = "Ex118 (excl. zeros) — Beta Fit", 
     xlab = "Score")

curve(dbeta(x, b_118["alpha"], b_118["beta"]), 
      add = TRUE, 
      col = "darkblue", 
      lwd = 2)

legend("topleft",
       legend = paste0("alpha=", round(b_118["alpha"],4),
                       "  beta=", round(b_118["beta"],4)),
       col = "darkblue", 
       lwd = 2, 
       bty = "n", 
       cex = 0.85)

# ExMA plot
hist(xMA_pos, 
     probability = TRUE, 
     breaks = 20, 
     col = "lightgreen",
     main = "ExMA (excl. zeros) — Beta Fit", 
     xlab = "Score")

curve(dbeta(x, b_MA["alpha"], b_MA["beta"]),
      add = TRUE, 
      col = "darkgreen", 
      lwd = 2)

legend("topleft",
       legend = paste0("alpha=", round(b_MA["alpha"],4),
                       "  beta=", round(b_MA["beta"],4)),
       col = "darkgreen", 
       lwd = 2, 
       bty = "n", 
       cex = 0.85)


# Beta Q-Q plots 
# Determining how good of a model Beta Dist. is 

par(mfrow = c(1, 2))

# Ex118
p118 <- ppoints(length(x118_pos))
q_th_118 <- qbeta(p118, b_118["alpha"], b_118["beta"])
q_em_118 <- sort(x118_pos)
plot(q_th_118, q_em_118,
     main = "Beta Q-Q: Ex118 (excl. zeros)",
     xlab = "Theoretical Quantiles (Beta)",
     ylab = "Sample Quantiles",
     pch = 16, cex = 0.6, col = "darkblue")
abline(0, 1, col = "red", lwd = 2)

# ExMA
pMA  <- ppoints(length(xMA_pos))
q_th_MA <- qbeta(pMA, b_MA["alpha"], b_MA["beta"])
q_em_MA <- sort(xMA_pos)
plot(q_th_MA, q_em_MA,
     main = "Beta Q-Q: ExMA (excl. zeros)",
     xlab = "Theoretical Quantiles (Beta)",
     ylab = "Sample Quantiles",
     pch = 16, cex = 0.6, col = "darkgreen")
abline(0, 1, col = "red", lwd = 2)

```


As seen in the Q-Q plots above, the beta models both datasets well, only deviating slightly at the lower tails (particularly for the ExMA dataset). As expected, due to the constraint from [0,1] on the Beta Distribution, it models the dataset much better than Gamma for values at the tails.

## 6. Model fit comparison using likelihood

#### Normal Distribution

Let $X_1,\ldots,X_n$ be IID such that:
$$
X_i \sim N(\mu,\sigma^2)
$$

The pdf is:

$$
f(x) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left(-\frac{(x-\mu)^2}{2\sigma^2}\right)
$$

Likelihood is given by:

$$
L(\mu,\sigma^2) = \prod_{i=1}^n f(x_i)=(2\pi\sigma^2)^{-n/2}exp\left(\frac{-1}{2\sigma^2}\sum_{i=1}^n (x_i-\mu)^2\right)
$$
Log-likelihood:

$$
\ell(\mu,\sigma^2)=\sum_{i=1}^n \log f(x_i) 
$$
$$
= -\frac{n}{2}\log(2\pi)-\frac{n}{2}\log(\sigma^2)-\frac{1}{2\sigma^2}\sum_{i=1}^n (x_i-\mu)^2
$$

#### Gamma Distribution

Let $X_1,....,X_n$ be IID such that: 
$$
X_i \sim \Gamma(\alpha,\beta)
$$
The pdf is:
$$
f(x)=\frac{\beta^\alpha}{\Gamma(\alpha)}x^{\alpha-1}e^{-\beta x}
$$
Likelihood given by:
$$
L(\alpha,\beta)=\prod_{i=1}^n\frac{\beta^\alpha}{\Gamma(\alpha)}x_i^{\alpha-1}e^{-\beta x_i}
$$
And grouping the terms gives: 
$$
L(\alpha,\beta)=\left(\frac{\beta^\alpha}{\Gamma(\alpha)}\right)^n\prod_{i=1}^nx_i^{\alpha-1}e^{-\beta \sum_{x} (x_i)}
$$

Log-likelihood:
$$
\ell(\alpha,\beta)=n\alpha\log\beta-n\log\Gamma(\alpha)+(\alpha-1)\sum_{i=1}^n\log x_i-\beta\sum_{i=1}^n x_i
$$

#### Beta Distribution

Let $X_1,....,X_n$ be IID such that:
$$
X_i \sim \beta(\alpha,\beta)
$$
The pdf is:
$$
f(x)=\frac{1}{B(\alpha,\beta)}x^{\alpha-1}(1-x)^{\beta-1}
$$
Likelihood:
$$
L(\alpha,\beta)=\prod_{i=1}^n\frac{1}{B(\alpha,\beta)}x_i^{\alpha-1}(1-x_i)^{\beta-1}
$$
Simplified to get: 
$$
L(\alpha,\beta)=
\left(\frac{1}{B(\alpha,\beta)}\right)^n
\prod_{i=1}^n
x_i^{\alpha-1}
\prod_{i=1}^n
(1-x_i)^{\beta-1}
$$

Log-likelihood:
$$
\ell(\alpha,\beta)= 
-n\log B(\alpha,\beta)
+(\alpha-1)\sum_{i=1}^n \log x_i
+(\beta-1)\sum_{i=1}^n \log(1-x_i)
$$

```{r}

loglik_norm1  <- sum(dnorm(x118_pos, mu118p, sig118p, log=TRUE))
loglik_norm2  <- sum(dnorm(xMA_pos, mu118p, sig118p, log=TRUE))

loglik_gamma1 <- sum(dgamma(x118_pos, 
                           round(mom_gamma(x118_pos)["alpha"],4), 
                           round(mom_gamma(x118_pos)["beta"],4), 
                           log=TRUE))

loglik_gamma2 <- sum(dgamma(xMA_pos, 
                           round(mom_gamma(x118_pos)["alpha"],4), 
                           round(mom_gamma(x118_pos)["beta"],4), 
                           log=TRUE))

loglik_beta1  <- sum(dbeta(x118_pos, 
                          round(mom_beta(x118_pos)["alpha"],4), 
                          round(mom_beta(x118_pos)["beta"],4), 
                          log=TRUE))

loglik_beta2  <- sum(dbeta(xMA_pos, 
                          round(mom_beta(x118_pos)["alpha"],4), 
                          round(mom_beta(x118_pos)["beta"],4), 
                          log=TRUE))

table <- matrix(c(loglik_norm1, 
                  loglik_norm2, 
                  loglik_gamma1, 
                  loglik_gamma2,
                  loglik_beta1, 
                  loglik_beta2), 
                nrow = 2) 

colnames(table) <- c("Normal", "Gamma", "Beta")
rownames(table) <- c("Ex118", "ExMA")

table
```
The best fitting model is determined by comparing log-likelihood values, where larger values indicate a better fit. For both datasets, the Normal Distribution has the highest log-likelihood, indicating that it provides the best fit among the three distributions considered.

